---
title: "Titanic"
author: "Frank D'Ambrosio"
date: "4/22/2021"
output: html_document
---
Variables
Survived (0,1)
Pclass (1,2,3)
Name
Sex (male, female)
Age 
SibSp 
Parch
Ticket
Fare
Cabin
Embarked


What I want to do:
fill in missing data
determine which variable are important
create a tree for each
create a forest from them
predict data with forest
write into final csv

Read in data
```{r}
sub <- read.csv("gender_submission.csv")
train <- read.csv("train.csv")
test <- read.csv("test.csv")
str(test)

(lived <- sum(train$Survived)/nrow(train))

#show if data is test or train
train$IsTrain <- TRUE
test$IsTrain <- FALSE


#need a survived column in test
test$Survived <- NA

names(train)
names(test)
ncol(train)
ncol(test)
#combine them
comb <- rbind(train , test)
names(comb)




(died <- 1-lived)

```

Load packages
```{r}
library(ggplot2) 
#library(ggthemes) 
#library(scales) 
library(dplyr) 
library(mice) 
library(randomForest) 
library(VIM)
```


factors
```{r}
train$Survived <- as.factor(train$Survived)
train$Pclass <- as.factor(train$Pclass)
train$Embarked <- as.factor(train$Embarked)




comb$Survived <- as.factor(comb$Survived)
comb$Pclass <- as.factor(comb$Pclass)
#comb$Age <- as.factor(comb$Age)
#comb$Ticket <- as.factor(comb$Ticket)
comb$Embarked <- as.factor(comb$Embarked)

```

Names/Titles
```{r}
# Grab title from passenger names
#https://www.journaldev.com/43690/sub-and-gsub-function-r
comb$Title <- gsub('(.*, )|(\\..*)', '', comb$Name)


(unique(comb[c("Title")]))
#We have some similar titles just spelled differently so we need to combine those
#Mr, Mrs, Miss, Master, Don, Rev, Dr, Mme, Ms, Major, Lady, Sir, Mlle, Col, Capt, the Countess, Jonkheer, Dona (only in test)
comb$Title[comb$Title == 'Mlle']<- 'Miss' #Mlle is Mademoiselle which is the unmarried version of Madame
comb$Title[comb$Title == 'Ms']<- 'Miss'
comb$Title[comb$Title == 'Mme']<- 'Mrs'
comb$Title[comb$Title == 'Capt' ] <- 'Nob'#for royalty/nobility/esteem
comb$Title[comb$Title == 'Col' ] <- 'Nob'
comb$Title[comb$Title == 'Don' ] <- 'Nob'
comb$Title[comb$Title == 'Dr' ] <- 'Nob'
comb$Title[comb$Title == 'Jonkheer' ] <- 'Nob'
comb$Title[comb$Title == 'Major' ] <- 'Nob'
comb$Title[comb$Title == 'Rev' ] <- 'Nob'
comb$Title[comb$Title == 'Sir' ] <- 'Nob'
comb$Title[comb$Title == 'the Countess' ] <- 'Nob'
comb$Title[comb$Title == 'Lady' ] <- 'Nob'
comb$Title[comb$Title == 'Dona' ] <- 'Nob'

comb$Title <- as.factor(comb$Title)

# Show title counts by sex
table(comb$Sex, comb$Title)

# Create a family size variable including the passenger themselves
comb$Fam_size <- comb$SibSp + comb$Parch + 1



```



Age
```{r}

age.out <- boxplot.stats(comb$Age)$stats[5] #looks like anything over 66 is a statistical outlier
age.out.filt <-  comb$Age < age.out #filter out ages over 66
comb[age.out.filt,]

age.equation <- "Age ~ Pclass + Sex + Fam_size + Embarked"

```


Missing Values
```{r}

aggr(comb)
matrixplot(comb)

colSums(is.na(train)) #177 missing values for age
colSums(is.na(test)) #86 age values missing in test, 1 fare missing, I don think that is significant
colSums(is.na(comb)) #263 for all which checks out




#Running MICE
summary(comb)
#mean age is 29.88
#mean Parch is .385
#mean Fare is 33.295
comb2 <- comb #create a second combined data set that will still have missingness

#age.comb.miss <- mice(comb, m = m, method = "cart")
colSums(is.na(comb))

#comb$Embarked[which(is.na(comb$Embarked))] = mode(comb$Embarked)
#comb$Age[which(is.na(comb$Age))] = mean(comb$Age, na.rm = TRUE) #fill in missing age
#comb$Age
comb$Fare[which(is.na(comb$Fare))] = mean(comb$Fare, na.rm = TRUE) #fill in missing fare

# want to look at age
#want to look at outliers of age
aggr(comb)
boxplot(comb$Age)


summary(comb2$Age)


#Mice Imputation 
#names(train)
#names(test)

comb2$Survived <- as.factor(comb2$Survived) #need to have these as factors 
comb2$Pclass <- as.ordered(comb2$Pclass)
comb2$Embarked <- as.factor(comb2$Embarked)

m = 5
titanic_imp <- mice(comb , m=5 , method = "cart" , maxit = 20 )
summary(titanic_imp)

imp_data <- complete(titanic_imp)

summary(comb$Age)


#now I want to take the imputed complete data set and use the complete variable columns and put them back into the comb2 data set

aggr(comb2) #here im just confirming the data is still missing

comb2$Age <- imp_data$Age
comb2$Pclass <- imp_data$Pclass
comb2$Embarked <- imp_data$Embarked
comb2$Fare[which(is.na(comb2$Fare))] <-  mean(comb$Fare[fare.out.filt], na.rm = TRUE)
#I also need to put the titles back in from the early transformation
comb2$Title <- comb$Title
summary(comb2)
#Kachow


```

Now I want to separate back into train and test sets
```{r}
t.train <- comb2[comb2$IsTrain == TRUE,]
t.test <- comb2[comb2$IsTrain == FALSE,]
summary(t.train)
summary(t.test)

t.train$Survived <- as.factor(t.train$Survived)
#str(t.train)


```


plots showing relationships in the training data set
```{r}
table(t.train$Survived, t.train$Age)

ggplot(aes(y = Age, x = Survived), data = t.train) + geom_boxplot()
ggplot(aes(y = Parch, x = Survived), data = t.train) + geom_boxplot()
ggplot(aes(y = Pclass, x = Survived), data = t.train) + geom_boxplot()
ggplot(aes(y = Fam_size, x = Survived), data = t.train) + geom_boxplot()
ggplot(aes(y = Embarked, x = Survived), data = t.train) + geom_boxplot()
ggplot(aes(y = Title, x = Survived), data = t.train) + geom_boxplot()
ggplot(aes(y = Fare, x = Survived), data = t.train) + geom_boxplot()

```

Now I want to create random forest based on the data
base model
```{r}

#Predict Survived
set.seed(7492)
ind <- sample(891, 623)
train_cv <- t.train[ind,]
test_cv <- t.train[-ind,]

str(train_cv)
#going to create the random forest here
# Pclass + Sex + Age + Fare + Embarked + Fam_size + Title
titanic.rf <- randomForest(as.factor(Survived)~ Pclass + Sex + Age + Fare + Embarked + Fam_size + Title , data= train_cv, ntree = 300 )
summary(test_cv)
plot(titanic.rf)

Survived <-  predict(titanic.rf , t.test)
PassengerId <- t.test$PassengerId
summary(Survived)
final <- data.frame(PassengerId, Survived)
write.csv(final, file = "Frank_Titanic_Predictions6.csv", row.names = FALSE)

```
model 6
```{r}


#Predict Survived
set.seed(7492)
ind <- sample(891, 623)
train_cv <- t.train[ind,]
test_cv <- t.train[-ind,]

str(train_cv)
#going to create the random forest here
# Pclass + Sex + Age + Fare + Embarked + Fam_size + Title
titanic.rf <- randomForest(as.factor(Survived)~ Pclass + Sex + Age + Fare + Embarked + Fam_size + Title , data= train_cv, ntree = 300 )
summary(test_cv)
plot(titanic.rf)

Survived <-  predict(titanic.rf , t.test)
PassengerId <- t.test$PassengerId
summary(Survived)
final <- data.frame(PassengerId, Survived)
write.csv(final, file = "Frank_Titanic_Predictions6.csv", row.names = FALSE)

```
model 7
```{r}

#Predict Survived
set.seed(7492)
ind <- sample(891, 623)
train_cv <- t.train[ind,]
test_cv <- t.train[-ind,]

str(train_cv)
#going to create the random forest here
# Pclass + Sex + Age + Fare + Embarked + Fam_size + Title
titanic.rf <- randomForest(as.factor(Survived)~ Pclass + Sex + Age  + Fam_size + Title , data= train_cv, ntree = 300 )
summary(test_cv)
plot(titanic.rf)

Survived <-  predict(titanic.rf , t.test)
PassengerId <- t.test$PassengerId
summary(Survived)
final <- data.frame(PassengerId, Survived)
write.csv(final, file = "Frank_Titanic_Predictions7.csv", row.names = FALSE)

```
model 8
```{r}

#Predict Survived
set.seed(7492)
ind <- sample(891, 623)
train_cv <- t.train[ind,]
test_cv <- t.train[-ind,]

str(train_cv)
#going to create the random forest here
# Pclass + Sex + Age + Fare + Embarked + Fam_size + Title
titanic.rf <- randomForest(as.factor(Survived)~ Pclass + Sex + Age + Fare + Embarked + Fam_size + Title , data= train_cv, ntree = 300 )
summary(test_cv)
plot(titanic.rf)

Survived <-  predict(titanic.rf , t.test)
PassengerId <- t.test$PassengerId
summary(Survived)
final <- data.frame(PassengerId, Survived)
write.csv(final, file = "Frank_Titanic_Predictions6.csv", row.names = FALSE)

```